---
title: "LISTA 5"
subtitle: "ANÁLISE MULTIVARIADA 1"
author: "Tailine J. S. Nonato"
date: "11/15/2023"
format: html
---


```{r}
#| message: false
#| echo: false
pacman::p_load(tidyverse)
```

### Exercício 9.1
Show that the covariance matrix
$$\rho = \begin{bmatrix}
 1 & .63 & .45\\
 .63 & 1 & .35\\
 .45 & .35 & 1
\end{bmatrix}
$$

for the $\rho=3$ standardized random variable $Z_{1}$,  $Z_{2}$ e $Z_{3}$ can be generated by the $m=1$ factor model

$$Z_{1} = .9F_{1} + \epsilon_{1}$$
$$Z_{2} = .7F_{1} + \epsilon_{2}$$
$$Z_{3} = .5F_{1} + \epsilon_{3}$$ 

where $Var(F_{1})=1$, $Cov(\epsilon,F_{1})=0$, and

$$Cov(\epsilon) = \begin{bmatrix}
 .19 & 0 & 0\\
 0 & .51 & 0\\
 0 & 0 & .75
\end{bmatrix}
$$

That is, write $ \rho $ in the form $ \rho = LL' + Cov(\epsilon) $

- Respostas

```{r}
lt <- matrix(c(.9,.7,.5),1,3)
lt
l <- t(lt)
l
llt <- l%*%lt
llt
cove<- matrix(c(.19,0,0,0,.51,0,0,0,.75),3,3)
rho<- llt + cove
rho
```

### Exercício 9.2
Use the information in 9.1

a. Calculate communalities $h_{i}^{2}$ for $i=1,2,3$, interpret these quantities.

- Respostas

Para $m=1$,

$h_{1}^{2}$ = $l_{11}^{2}$ = `r l[1,1]^2`

$h_{2}^{2}$ = $l_{21}^{2}$ = `r l[2,1]^2`

$h_{3}^{2}$ = $l_{31}^{2}$ = `r l[3,1]^2`

Communalities: Partes das variâncias explicadas pelo fator único.

b. Calculate $Corr(Z_{i},F_{1})$ for $i=1,2,3$. Which variable might carry the greatest weight in "naming" the common factor? Why?

- Respostas

$Corr(Z_{1},F_{1})$ = $l_{11}$ = `r l[1,1]`

$Corr(Z_{2},F_{1})$ = $l_{21}$ = `r l[2,1]`

$Corr(Z_{3},F_{1})$ = $l_{31}$ = `r l[3,1]`


$Z_{1}$ tem a maior correlação com o fator, dessa forma será a variável que mais carregará o "nome" do fator.

### Exercício 9.3

```{r}

```