---
title: "Lista 1"
subtitle: "ESTATISTICA COMPUTACIONAL"
author: "Tailine J. S. Nonato"
date: "12/09/2023"
format:
  html:
    code-fold: true
---

## Rizzo 5.2

Q: Compute a Monte Carlo estimate of the standard
normal cdf, by generating from the Uniform($\theta$,$x$) distribution. Compare your estimates with the normal cdf function pnorm. Compute an estimate of the variance of your Monte Carlo estimate of $\phi$(2), and a 95% confidence interval for $\phi$(2).

```{r, echo=FALSE, message=FALSE}
#loading packages
if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse)
```

To answer the questions, the first step is creating the 

```{r}
#creating a function
phi.mc <- function(x, n=1000){
    x.abs=abs(x)
    i=(1/n)*(x.abs/sqrt(2*pi))*sum(exp(((-runif(n,max=x.abs)^2)/2)))
    if(x>=0){return(0.5+i)}
    if(x<0){return(0.5-i)}
}
```

If we run the function using x=2 and n=1000, we get:
```{r}
phi.mc(2,1000)
```

Now, comparing the estimates with pnorm we get the following table:

```{r}
#comparing with normal dist
x.seq<-seq(-3,3,0.5)
phi.mc.seq<- map(x.seq, phi.mc)
pnorm.seq<-pnorm(x.seq)
cbind(x.seq,phi.mc.seq,pnorm.seq)
```

Now, estimating the variance of the Monte Carlo estimate of $\phi(2)$, and a $95%$ confidence interval
for $\phi(2)$, we'd have:

r=100

phi.mc.2<- function()phi.mc(2,1000)

est.phi<-map(1:r,phi.mc.2)

quantile(est.phi,c(0.0025,0.975))


